Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system with plausible mechanism

### 2024-07-19 _tex/index.tex
Built site for gh-pages

@@

@@

@@

@@

\subsection{Why estimate judges appart?}\label{sec-appA}

>>>>

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{images/figures/ACflow.png}

}

\caption{\label{fig-ACflow}The flow of association and causation in
graphs. Extracted from \citet[31]{Neal_2020}}

\end{figure}%

@@

\begin{figure}

\centering{

\includegraphics[width=0.4\textwidth,height=\textheight]{images/figures/IEflow.png}

}

\caption{\label{fig-IEflow}Identification-Estimation flowchart.
Extracted from \citet[32]{Neal_2020}}

\end{figure}%

@@

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{images/figures/SciModel_simp1.png}

}

\caption{\label{fig-SciModel_simp1}DCJ causal diagram, simplified
description}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{images/figures/SciModel_simp2.png}

}

\caption{\label{fig-SciModel_simp2}DCJ causal diagram, simplified
mathematical description}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{images/figures/SciModel_pop.png}

}

\caption{\label{fig-SciModel_pop}DCJ causal diagram, population
mathematical description}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{images/figures/SciModel_samp.png}

}

\caption{\label{fig-SciModel_samp}DCJ causal diagram, sample with
comparisons mathematical description}

\end{figure}%

\subsection{Probabilitics assumptions of the scientific
model}\label{sec-theory-probability}

\newcommand{\dsep}{\perp\!\!\!\perp}
\newcommand{\ndsep}{\not\!\perp\!\!\!\perp}

\begin{equation}\phantomsection\label{eq-StructuralModel}{
\begin{aligned}
  O_{kijr_{ij}} := & \; f_{O}( \delta_{kijr_{ij}} ) \\
  \delta_{kijr_{ij}} := & \; f_{D}( \gamma_{k}, \theta_{ir_{i}} ) \\
  \gamma_{k} := & \; f_{G}( Z_{k}, e_{k} ) \\
  \theta_{ir_{i}} := & \; f_{R}( \theta_{i}, Y_{r}, e_{r_{i}} ) \\
  \theta_{i} := & \; f_{T}( X_{i}, e_{i} ) \\
  & \; e_{k} \perp\!\!\!\perp e_{i} \\
  & \; e_{k} \perp\!\!\!\perp e_{r_{i}} \\
  & \; e_{i} \perp\!\!\!\perp e_{r_{i}}
\end{aligned}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-StatModel_general}{
\begin{aligned}
  O_{kijr_{ij}} \sim & \; \text{Bernoulli}\left[ \; logit^{-1}\left( \delta_{kijr_{ij}} \right) \; \right] \\
  \delta_{kijr_{ij}} = & \; \gamma_{k}( \theta_{ir_{i}} - \theta_{jr_{j}} ) \\
  \gamma_{k} = & \; logit^{-1}\left[ \beta_{Z} Z_{k} + e_{k} \right] \\
  \theta_{ir_{i}} = & \; \theta_{i} + \beta_{Y} Y_{r} + e_{r_{i}}  \\
  \theta_{i} = & \; \beta_{X} X_{i} + e_{i}  \\
  e_{k} \sim & \; \text{Normal}(0, \sigma_{k(g)}) \\
  e_{i} \sim & \; \text{Normal}(0, \sigma_{i(g)}) \\
  e_{r_{i}} \sim & \; \text{Normal}(0, \sigma_{r(g)})
\end{aligned}
}\end{equation}

for identification purposes we can set
\(\frac{1}{G}\sum_{g=1}^{G} \sigma_{k(g)} = 0.02\),
\(\frac{1}{G}\sum_{g=1}^{G} \sigma_{i(g)} = 1\), and
\(\frac{1}{G}\sum_{g=1}^{G} \sigma_{r(g)} = 1\). A special case of this
would be to assume that the data comes from the same population, in that
case, \(\sigma_{k(g)} = \sigma_{k} = 0.02\),
\(\sigma_{i(g)} = \sigma_{i} = 1\)

\subsection{Let's talk about Thurstone}\label{sec-theory-thurstone}

Thurstone's comparative judgment \citet{Thurstone_1927} is based on the
formula: \[
X_{AB} = \frac{S_{A} - S_{B}}{\sigma_{AB}}
\]

where \(X_{AB}\) defines the comparative judgment outcome, \(S_{A}\) and
\(S_{B}\) are the modal discriminal processes,
\(\sigma_{AB} = \sqrt{ \sigma^{2}_{A} + \sigma^{2}_{B} + 2 \rho \sigma_{A} \sigma_{B}}\),
with \(\sigma_{A}\) and \(\sigma_{B}\) being the dispersion of
discriminal processes \(A\) and \(B\), respectively, and \(\rho\) the
correlation between discriminal processes.

The theory identifies five cases:

\begin{itemize}
\tightlist
\item
  \textbf{Case 1:} only constant \(\rho\) (not \(\rho_{ij}\))
\item
  \textbf{Case 2:} \(X_{ij}\) becomes \(X_{kij}\) with \(k=1, \dots, K\)
  judges (replication, not duplication)
\item
  \textbf{Case 3:} \(\rho = 0\), then
  \(\sigma_{AB} = \sqrt{ \sigma^{2}_{A} + \sigma^{2}_{B}}\)
\item
  \textbf{Case 4:} \(\sigma_{B}=\sigma_{A}+d\), then
  \(\lim_{d \leq 0.1\sigma_{A}} \sigma_{AB} = (\sigma_{A} + \sigma_{B}) /\sqrt{2}\)
\item
  \textbf{Case 5:} \(\sigma_{B}=\sigma_{A}\), then
  \(\sigma_{AB} = \sqrt{2}\sigma\)
\end{itemize}

Now using the DAG and statistical notation

\begin{equation}\phantomsection\label{eq-StatModel_thurstone}{
\begin{aligned}
  O_{kijr_{ij}} := & \; f_{O}( \delta_{kijr_{ij}} ) \\
  \delta_{kijr_{ij}} = & \; \gamma_{k}( \theta_{ir_{i}} - \theta_{jr_{j}} ) \\
  \gamma_{k} = & \; f_{G}( Z_{k}, e_{k} ) \\
  \theta_{ir_{i}} = & \; \theta_{i} + \beta_{Y} Y_{r} + e_{r_{i}}  \\
  \theta_{i} = & \; \beta_{X} X_{i} + e_{i}  \\
  e_{k} \sim & \; \text{Normal}(0, \sigma_{k(g)}) \\
  e_{i} \sim & \; \text{Normal}(0, \sigma_{i(g)}) \\
  e_{r_{i}} \sim & \; \text{Normal}(0, \sigma_{r(g)})
\end{aligned}
}\end{equation}

The theory identifies five cases:

\begin{itemize}
\tightlist
\item
  \textbf{Case 1:} only constant \(\rho \approx \sigma_{i}\)
\item
  \textbf{Case 2:} now judges are separated by using \(\gamma_{k}\)
\item
  \textbf{Case 3:} \(\rho \approx \sigma_{e_{i}} = 0\) (no nesting of
  texts on students), then
  \(\sigma_{AB} = \sqrt{ \sigma^{2}_{A} + \sigma^{2}_{B}}\)
\item
  \textbf{Case 4:} \(\sigma_{B}=\sigma_{A}+d\), then
  \(\lim_{d \leq 0.1\sigma_{A}} \sigma_{AB} = (\sigma_{A} + \sigma_{B}) /\sqrt{2}\)
\item
  \textbf{Case 5:} \(\sigma_{B}=\sigma_{A}\), then
  \(\sigma_{AB} = \sqrt{2}\sigma\)
\end{itemize}

But now can we see other scenarios than just those 5 cases?

\begin{itemize}
\tightlist
\item
  consider different \(\rho \approx \sum_{p=1}^{P} \sigma_{p}\),
  depending on \(P\) nesting structures
\item
  we can now investigate \(\gamma_{k}\)
\item
  we can assume \(\sigma_{B} \neq \sigma_{A}\), no need for results on
  the limit
\end{itemize}

@@

\subsection{Why do we need to estimate judges'
abilities?}\label{sec-appA}
